---
title: "Motor Trends"
author: "Ryan P"
date: "5/07/2021"
output: pdf_document
---

# Question One

- "Is an automatic or manual transmission better for MPG"

#### Exploratory Analyses


```{r, include = FALSE}
library(ggplot2)
library(dplyr)
```


A scatter plot was used at first (See Appendix), but a violin plot best visualized the data quickly and showed that **a manual transmission is better for MPG**.

```{r, fig.width = 5, fig.height = 3}
cars <- arrange(mtcars, am)
g <- ggplot(cars, aes(x = 1:32, y = mpg, col = as.factor(am)))
g + geom_violin()
```

The plot does not indicate if the difference is significant or if the difference in fuel efficiency is due to the transmission type.

A linear regression model was then used to see if there was significance and how much effect there was on fuel efficiency.

```{r}
mfit <- lm(mpg ~ am, mtcars)
summary(mfit)$coef 
```
# Question Two

- "Quantify the MPG difference between automatic and manual transmissions"


From the coefficients, we can see that the average fuel efficiency for automatic transmissions is 17.147 mpg. The second coeffiecient in the "am" row is **the difference in fuel efficiency of 7.244 mpg** that is apparent between manual and automatic transmissions. The average manual car therefore has an mpg of 24.392.
The R^2 value of `r summary(mfit)$r.squared`, indicates that the transmission factor can only explain 36% of the variation.

## Executive Summary

Though there appears to be a distinct improvement in gas mileage in manual vehicles,
further analysis suggests that it is not due to the transmission, but to other
confounding factors such as weight, cylinders, and displacement (See Appendix).


## ANALYZE the Residual Plots
Looking at the residual plots of fuel efficiency vs transmission model produces:

```{r fig.cap= "Transmission Model"}
par(mfrow = c(2,2))
plot(mfit)
```
The first plot is heteroskedastic in that it shows non-constant variance and a spread in the residuals with the different transmissions.
The Q-Q plot does indicate that the residuals are normally distributed.
The Scale-Location reflects the spread of the residuals with a non-horizontal line.
The last plot does not make it easy to identify any outliers or influencers.

# APPENDIX

# Further Linear Regressions

ANOVA was used to identify other significant alternatives to transmission for causing variation in fuel efficiency. The linear regression including these factors resulted in higher R-squared values.
```{r}

allfit <- lm(mpg ~., mtcars)
anova(allfit)
```
Cylinders, Displacement, and Weight come out as significant and worth exploring.

```{r}
bitfit <- lm(mpg~ cyl + disp+ wt, mtcars)
summary(bitfit)$coef
```
The linear regression of fuel efficiency due to cylinders, displacement, and weight has an R^2 value of `r summary(bitfit)$r.squared` which is closer to that of the model including all variables (R-squared =`r summary(allfit)$r.squared`)


Residual plots for the model involving these three variables show better homoskedasticity than the transmission model.
```{r, fig.cap = "Three Factor Model"}
par(mfrow = c(2,2))
plot(bitfit)
```


```{r Scatter, fig.width = 5, fig.height = 3, echo = FALSE, fig.cap = "Scatter Plot"}
cars <- arrange(mtcars, am)
g <- ggplot(cars, aes(x = 1:32, y = mpg, col = as.factor(am)))
g + geom_point()
```

